Re-fit

### 2023-01-16 scripts/r/04_refit/refit_tr.Rmd
🐛 - fix procambarus_maculosus

# libs
library(tidyverse)
library(brms)
library(readxl)
library(lme4)
library(here)

@@

data <- read.delim("./data/analyses/tr/selectedData.txt", header = T, sep = "") %>% 
  as_tibble() # %>% 
  # filter(sex == "female") # only run on female data!

data = read.csv("./data/analyses/tr/pervagor_meeki/Data/ManySpeech_Praat_merged_data.csv", header=T)

info <- data %>% 
  select(sl_rowIdx, bundle, vowel, speaker, sex, typicality, word) %>% 
  unique()

@@

# create a basis for B-splines
basis <- create.bspline.basis(rangeval = c(0, 1), nbasis = nBasis, norder = nOrder)

# define functional parameter object (will be used to estimate functional parameters later)
fdParObj <- fdPar(fdobj = basis, Lfdobj = Lfdobj, lambda = lambda)

coef <- data %>%
  group_by(sl_rowIdx) %>%
  summarise(coef = c(smooth.basis(argvals = times_norm, y = F0, fdParObj)$fd$coefs),
            coefId = seq_len(nBasis)) %>%
  ungroup() %>%
  pivot_wider(names_from = coefId, values_from = coef) %>%
  select(-sl_rowIdx) %>% 
  abind(along = 2) %>%
  aperm(c(2,1))

@@

# create functional data object
y_fd <- fd(coef = coef, basisobj = basis)

# this is the actual FPCA
nPCs <- 3
y_pcafd <- pca.fd(y_fd, nharm = nPCs, fdParObj)
# saveRDS(y_pcafd, file = "./pcafd.rds")

@@

curves <- expand.grid(time = tx,
                      PC = 1:3,
                      perturbation = c(-1, 0, 1)
) %>% setDT

@@

curves[, value := (y_pcafd$meanfd$coefs[,1] + perturbation * sd_score[PC] * y_pcafd$harmonics$coefs[,PC]) %>% 
         fd(., basis) %>%
         eval.fd(tx, .),
       by = .(PC, perturbation)]

@@

pl <- ggplot(reconstruct) + 
  aes(x = time, y = value) + 
  geom_line(size = 1.2) + 
  facet_grid(k~per, labeller = label_parsed) +
  xlab("Normalised Time") + ylab("F0 [Hz]") + theme + 
  guides(linetype = guide_legend(override.aes = list(size = 1)))
ggsave(filename = "./PCcurves.png", plot = pl, width = 25, height = 20, unit = "cm")

pl <- ggplot(info %>% pivot_longer(cols = s1:s3) %>% 
               mutate(score = case_when(name == "s1" ~ "s[1]",
                                        name == "s2" ~ "s[2]",
                                        name == "s3" ~ "s[3]"))) + 
  aes(x = 0, y = value, fill = typicality) + 
  facet_grid(score~., scales = "free_y", labeller = label_parsed, switch = "y") + 
  geom_boxplot() + ylab("") + xlab("") +
  theme + theme(strip.background.y = element_rect(fill = "white"), 
                strip.placement = "outside",
                legend.margin = margin(t = -15))
ggsave(filename = "./PCscores.png", plot = pl, width = 18, height = 18, unit = "cm")

@@

curves <- expand.grid(time = tx,
                      PC = 1:2,
                      perturbation = seq(-1, 1, by = .25)
) %>% setDT

@@

pl <- ggplot(curves) +
  aes(x = time, y = value, col = perturbation, group = perturbation) +
  geom_line(size = 1.2) +
  scale_color_gradient2(name = "P ", low = "cornflowerblue", mid = "grey", high = "darkgoldenrod1") +
  facet_grid(~PC, labeller = PC_labeller) +
  geom_line(data = curves[perturbation == 0], color = "black", size = 1.5) +
  xlab("Normalised Time") + ylab("F0 [Hz]") +
  theme + theme(legend.direction = "horizontal", 
                legend.title = element_text(),
                legend.text = element_text(size = 12))
ggsave(filename = "./PCcurves_col.png", plot = pl, width = 20, height = 12, unit = "cm")

>>>>

# libs
library(tidyverse)
library(brms)
library(readxl)
library(lme4)
library(lmerTest)
library(here)

@@

data <- read.delim("./data/analyses/tr/procambarus_maculosus/selectedData.txt", header = T, sep = "") %>% 
  as_tibble() # %>% 
  # filter(sex == "female") # only run on female data!

# data = read.csv("./data/analyses/tr/pervagor_meeki/Data/ManySpeech_Praat_merged_data.csv", header=T)

info <- data %>% 
  dplyr::select(sl_rowIdx, bundle, vowel, speaker, sex, typicality, word) %>% 
  unique()

@@

# create a basis for B-splines
basis <- fda::create.bspline.basis(rangeval = c(0, 1), nbasis = nBasis, norder = nOrder)

# define functional parameter object (will be used to estimate functional parameters later)
fdParObj <- fda::fdPar(fdobj = basis, Lfdobj = Lfdobj, lambda = lambda)

coef <- data %>%
  group_by(sl_rowIdx) %>%
  summarise(coef = c(fda::smooth.basis(argvals = times_norm, y = F0, fdParObj)$fd$coefs),
            coefId = seq_len(nBasis)) %>%
  ungroup() %>%
  pivot_wider(names_from = coefId, values_from = coef) %>%
  select(-sl_rowIdx) %>% 
  abind::abind(along = 2) %>%
  aperm(c(2,1))

@@

# create functional data object
y_fd <- fda::fd(coef = coef, basisobj = basis)

# this is the actual FPCA
nPCs <- 3
y_pcafd <- fda::pca.fd(y_fd, nharm = nPCs, fdParObj)
# saveRDS(y_pcafd, file = "./pcafd.rds")

@@

curves <- expand.grid(time = tx,
                      PC = 1:3,
                      perturbation = c(-1, 0, 1)
) %>% data.table::setDT()

@@

curves[, value := (y_pcafd$meanfd$coefs[,1] + perturbation * sd_score[PC] * y_pcafd$harmonics$coefs[,PC]) %>% 
         fda::fd(., basis) %>%
         fda::eval.fd(tx, .),
       by = .(PC, perturbation)]

@@

# pl <- ggplot(reconstruct) + 
#   aes(x = time, y = value) + 
#   geom_line(size = 1.2) + 
#   facet_grid(k~per, labeller = label_parsed) +
#   xlab("Normalised Time") + ylab("F0 [Hz]") + theme + 
#   guides(linetype = guide_legend(override.aes = list(size = 1)))
# ggsave(filename = "./PCcurves.png", plot = pl, width = 25, height = 20, unit = "cm")
# 
# pl <- ggplot(info %>% pivot_longer(cols = s1:s3) %>% 
#                mutate(score = case_when(name == "s1" ~ "s[1]",
#                                         name == "s2" ~ "s[2]",
#                                         name == "s3" ~ "s[3]"))) + 
#   aes(x = 0, y = value, fill = typicality) + 
#   facet_grid(score~., scales = "free_y", labeller = label_parsed, switch = "y") + 
#   geom_boxplot() + ylab("") + xlab("") +
#   theme + theme(strip.background.y = element_rect(fill = "white"), 
#                 strip.placement = "outside",
#                 legend.margin = margin(t = -15))
# ggsave(filename = "./PCscores.png", plot = pl, width = 18, height = 18, unit = "cm")

@@

curves <- expand.grid(time = tx,
                      PC = 1:2,
                      perturbation = seq(-1, 1, by = .25)
) %>% data.table::setDT()

@@

# pl <- ggplot(curves) +
#   aes(x = time, y = value, col = perturbation, group = perturbation) +
#   geom_line(size = 1.2) +
#   scale_color_gradient2(name = "P ", low = "cornflowerblue", mid = "grey", high = "darkgoldenrod1") +
#   facet_grid(~PC, labeller = PC_labeller) +
#   geom_line(data = curves[perturbation == 0], color = "black", size = 1.5) +
#   xlab("Normalised Time") + ylab("F0 [Hz]") +
#   theme(legend.direction = "horizontal", 
#                 legend.title = element_text(),
#                 legend.text = element_text(size = 12))
# ggsave(filename = "./PCcurves_col.png", plot = pl, width = 20, height = 12, unit = "cm")

```{r procambarus_maculosus_models}

s1.lmer <- lmer(s1 ~ typicality * sex + (sex | word) + (typicality | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) %>% 
  step() %>% get_model()

# they did not interpret the models with typicality in there, so the minimum model we refit will be typicality + sex

s2.lmer <- lmer(s2 ~ typicality * sex + (sex | word) + (typicality | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) %>% 
  step() %>% get_model()
# s2 ~ typicality + (typicality | speaker), but failed to converge

s2.lmer <- lmer(s2 ~ typicality + (1 | word) + (1 | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) 
s2.lmer %>% summary(corr = F)

```{r procambarus_maculosus_refitting}

dat <- info %>% 
  mutate(
         # scale continuous variables
         s1_s = (s1 - mean(s1, na.rm = TRUE)) / sd(s1,  na.rm = TRUE),
         s2_s = (s2 - mean(s2, na.rm = TRUE)) / sd(s2,  na.rm = TRUE),
         sex_s = as.factor(sex),
         sex_s = C(sex_s, sum),
         effect_cat = as.factor(typicality),
         effect_cat = C(effect_cat, treatment)
  )

# run brms on standardized models
procambarus_maculosus_1_pca1_cat <- brm(
  s1_s ~ 
         effect_cat + sex_s + 
    (sex_s | word) + 
    (effect_cat | speaker),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_maculosus_1_pca1_cat"
  )

# run brms on standardized models
procambarus_maculosus_2_pca2_cat <- brm(
  s2_s ~ 
         effect_cat + 
    (1 | word) +
    (1 | speaker),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_maculosus_2_pca2_cat"
  )
